---
title: "Untitled"
author: "Haotian Bu"
date: "February 2, 2019"
output: html_document
---


```{r getData, eval=FALSE}
link='https://github.com/EvansDataScience/VisualAnalytics_2_tabularData/raw/master/data/eduwa.rda'

#getting the data TABLE from the file in the cloud:
load(file=url(link))
```


```{r freqTab_rel, eval=FALSE}
# relative values, proportion . the input of the prop is a table
absoluteT=table(eduwa$LocaleType,exclude = 'nothing')
prop.table(absoluteT)
```

This table tells us the location of the public schools. What is the right visual for this? Sometimes the answer seems obvious, as tradition or habits give so much weight to decisions. Let's use the very well known **pie chart**:

```{r pieBasic, eval=FALSE}
# the pie plots the table:
ToPlot=prop.table(absoluteT)
```

You should always keep it simple. Then decorate. For example, you can start improving the plot you already *have*:

We could alter the fifth label:
```{r alterTable, eval=FALSE}
names(ToPlot)[5]='Unknown'
```


* Our plot did not have a title. Titles (and subtitles) are important. To give a title, it can be a *question* to be answered by the plot:


```{r pieTitle_question, eval=FALSE}
# the pie plots the table:
titleText='Where are Public Schools located in WA in 2019?'
sourceText='Source: US Department of Education'
```

The title can *guide* the reader to recognise the purpose of your plot:

```{r pieTitle_guide, eval=FALSE}
# the pie plots the table:
titleText2='WA still has schools locations unknown \n (info from 2018)'
```


The title can also *suggest* the decision:

```{r pieTitle_suggestion, eval=FALSE}
# the pie plots the table:
titleText3='WA needs to fully categorize school locations\n(info from 2018)'
```


* You can customize the colors:

```{r rainbowColor, eval=FALSE}
# details:
ToPlot=ToPlot*100 # preparing labels
```

```{r ggplot1, eval=FALSE}

tableFreq=as.data.frame(ToPlot)
names(tableFreq)=c("locale","pct")
#you have:
tableFreq
```


```{r Excise1_BarPlotByShare, eval=FALSE}
library(ggplot2)

exercise1 <- ggplot(data = tableFreq, aes(x = reorder(locale, pct), y = pct)) +
  geom_bar(fill ="gray", stat = 'identity') +
  labs(title=titleText2,x =NULL, y = NULL, caption = sourceText) +
  geom_hline(yintercept = 20,linetype="dashed", size=1.5, alpha=0.5) +
  scale_y_continuous(breaks=c(0,10, 25,40,50),limits = c(0, 50), 
                     labels=scales::unit_format(suffix = '%')) +
  theme(panel.background = element_rect(fill = "white",colour = "grey50"),
        plot.caption = element_text(hjust = 0),
        plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(y = pct ,label = paste0(round(pct,2), '%')),
            hjust=1, size = 3) +
  coord_flip() +
  theme_bw()
  
exercise1
```


<span style="color:red"> Exercise:<br>Complete adding the elements missing in the last plot. </span>


```{r EXercise2_lolliplot}
tableFreqO=tableFreq[order(tableFreq$pct),]
tableFreqO$gap=tableFreqO$pct-25
tableFreqO$PositiveGap=ifelse(tableFreqO$gap>0,T,F)


ggplot(tableFreqO, aes(locale,gap,color=PositiveGap, label = round(gap,3))) +
  geom_segment(aes(y = 0, x = locale, yend = gap, xend = locale), color = "gray") + 
  geom_point() +
  scale_x_discrete(limits=tableFreqO$locale) +
  labs(title=titleText2,x =NULL, y = NULL, caption = sourceText) +
  geom_text(nudge_x=0.3,show.legend = FALSE) +
  theme_bw()

```

```{r Exercise3_dataset}
frqTabO=as.data.frame(prop.table(table(eduwa$High.Grade))) #ggplot deals with dataframe, transfer tables into dataframes
frqTabO
names(frqTabO)=c('grade','pct') # give names to variables
frqTabO
```

```{r Exercise3_BoxPlot}
titleText3 <-" The distribution of the highest grade offered in a school"

ordLabels= levels(eduwa$High.Grade)

ggplot(eduwa, aes(y=as.numeric(High.Grade))) + 
  geom_boxplot() + coord_flip() +# to show it horizontally, y-axis has no meaning 
  labs(title=titleText3,x =NULL, y = NULL, caption = sourceText) +
  scale_y_continuous(labels=ordLabels, breaks=1:15) +
  geom_hline(yintercept = mean(as.numeric(eduwa$High.Grade)),linetype="dashed", size=1.5, alpha=0.5) +
  theme(panel.background = element_rect(fill = "white",colour = "grey50"),
        panel.grid = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.caption = element_text(hjust = 0),
        plot.title = element_text(hjust = 0.5))
  
```

```{r Exercise4_HistogramFreeLunch}
eduwa5=eduwa[complete.cases(eduwa$Free.Lunch),]

TitleText4 <- "The distribution of students who have Free Lunch"

ggplot(eduwa5,aes(x = Free.Lunch)) + 
  geom_histogram(binwidth = 100,boundary=0) +#changing width
  stat_bin(binwidth = 100, aes(label=..count..), 
           geom = "text",boundary = 0,vjust=-0.5) +
  labs(title=TitleText4,x =NULL, y = NULL, caption = sourceText) +
  theme(plot.title = element_text(hjust = 0.5))
```

Exercise:<br> Do some research and make a histogram and a density plot using ggplot for the variable we just used above.

```{r Exercise5_Histogram}
eduwa4=eduwa[complete.cases(eduwa$Student.Teacher.Ratio),]
meanValues=tapply(eduwa4$Student.Teacher.Ratio, 
                  eduwa4$County, 
                  mean)
TableMean=as.data.frame(meanValues)

TitleText5 <- "Mean value of Student-Teacher Ratio" 

ggplot(TableMean,aes(x = TableMean$meanValues)) + 
  geom_histogram(binwidth = 2,boundary=0) +#changing width
  stat_bin(binwidth = 2, aes(label=..count..),
           geom = "text",boundary = 0,vjust=-0.5) +
  labs(title=TitleText5,x =NULL, y = NULL, caption = sourceText) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Exercise5_DensityPlot}


TitleText5 <- "Density plot of Student-Teacher Ratio"

ggplot(TableMean, ) +
  stat_density(aes(x = meanValues), adjust = 1, geom = "line", kernel = "gaussian") +
  geom_vline(xintercept = mean(meanValues), linetype="dashed", size=1.5, col = 'blue', alpha=0.5) +
  geom_vline(xintercept = median(meanValues),lty="dashed",lwd=1.5,col='red') +#median  
  labs(title=TitleText5, x =NULL, y = NULL, caption = sourceText) +
  theme(plot.title = element_text(hjust = 0.5))
```

