---
title: "Hw1"
author: "Haotian Bu"
date: "2019年1月16日"
output: html_document
---


So far, we have used the _basic_ R capabilities for plotting. 

There are alternative libraries, like **ggplot2**, that are also frequently used. However, it has a different approach, which allows to add layers that let you customize your plot. The classic approach for ggplot is:

* Avoid missing values and prepare frequency table. We replaced the missing values (now they are 'Unknown'). Here, you need to transform the _table_ into a **data frame**:

```{r ggplot1, eval=FALSE}

tableFreq=as.data.frame(ToPlot)
names(tableFreq)=c("locale","pct")
#you have:
tableFreq
```


* Call the library:

```{r ggplot_call, warning=FALSE}
library(ggplot2)
```

* Create the base _object_, which is not a plot, just informing the main variables:

```{r ggplot_base, eval=FALSE}
#base GGPLOT2 starts with a "base", telling WHAT VARIABLES TO PLOT
base= ggplot(data = tableFreq, 
             aes(x = locale,
                 y = pct)) 
base
```

* On top of the previous object, add the layer that produces the main plots (the next layers will add or customize elements in the plot):

```{r bar_gg, eval=FALSE}
plot1 = base + geom_bar(fill ="gray",
                        stat = 'identity') # y is just what it is!
plot1
```

* We can now pay attention to the titles:

```{r title_gg, eval=FALSE}
plot2 = plot1 + labs(title=titleText2,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText)
plot2

```

* Add the reference lines:

```{r refl_gg, eval=FALSE}
plot3 = plot2 + geom_hline(yintercept = 25, #where
                           linetype="dashed", 
                           size=1.5, #thickness
                           alpha=0.5) #transparency
plot3
```

* Customize the axes:

```{r units_gg, eval=FALSE}
library(scales)

# customize Y axis
plot4 = plot3 + scale_y_continuous(breaks=c(0,10, 25,40,50),
                                 limits = c(0, 50), # expand = c(0, 0),
                                 labels=scales::unit_format(suffix = '%')) 
plot4
```

* Less ink and title/subtitle positions:
```{r title2_gg, eval=FALSE}
plot5 = plot4 + theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.caption = element_text(hjust = 0), # default was 1
                    plot.title = element_text(hjust = 0.5))
plot5
```

* annotating the bars:

```{r annot_gg, eval=FALSE}
plot6 = plot5 + geom_text(aes(y = pct ,
                            label = paste0(round(pct,2), '%')),
                        vjust=1, # if flipping 'hjust'
                        size = 3)
# wanna flip the plot?
plot6 #+ coord_flip()

```


Bar plots are the default option for categorical variables. In general, you see the distribution of the classification, which allows you to identify **concentration**. For that reason, ordering the bars by height can be helpful:

```{r orderBars_base, eval=FALSE}
###
ToPlotOrd=sort(ToPlot)
###

par(mgp=c(1,0.5,0)) # distance label, tickText,tick
location=barplot(ToPlotOrd,
         col = paletteHereNew,
         border=NA,
         main=titleText2,
         xlim=c(0,50), 
         xlab = '(in %)',
         horiz = T,
         cex.names = 0.7,
         cex.lab=0.6,
         xaxt="n") # no x-axis, so I customize it below...

text(x=ToPlotOrd,y=location,labels=plotLabels,pos = 2,cex = 0.7) 
title(sub=sourceText, adj=0,cex.sub=0.7,line = 3) 

# reference line
abline(v=25,lty=3,lwd=3)

# customizong tick values
xtick<-c(0,10,25,40,50)
axis(side=1, at=xtick, labels = xtick,cex.axis=0.8)

```

The plot above simply change the order of the _table_. If you want to do the same with _ggplot_ you should try the command:

```{r orderFor_gg, eval=FALSE}

ggplot(data = tableFreqOrd, aes(x = reorder(locale, pct), y = pct)) +
  geom_bar(fill ="gray", stat = 'identity') +
  labs(title=titleText2,x =NULL, y = NULL, caption = sourceText) +
  geom_hline(yintercept = 20,linetype="dashed", size=1.5, alpha=0.5) +
  scale_y_continuous(breaks=c(0,10, 25,40,50),limits = c(0, 50), 
                     labels=scales::unit_format(suffix = '%')) +
  theme(panel.background = element_rect(fill = "white",colour = "grey50"),
        plot.caption = element_text(hjust = 0),
        plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(y = pct ,label = paste0(round(pct,2), '%')),
            hjust=1, size = 3) +
  coord_flip() +
  theme_bw()
  

```


<span style="color:red"> Exercise:<br>Use ggplot to show a bar plot ordered by share size.</span>



We could use our reference line to show gaps or differences. In this case, the **Lollipop** plot may be useful. This one is just a replacement for a bar plot:

```{r lolli_gg, eval=FALSE}
base = ggplot(tableFreq, aes(x=locale,pct)) 
lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = locale, 
                                   yend = pct, 
                                   xend = locale), color = "grey50") 
lolliplot1 + geom_point()
```

And, if you order the data frame:
```{r lolli_ord, eval=FALSE}
tableFreq[order(tableFreq$pct),]
```

You can get:
```{r lolli_ord_gg, eval=FALSE}

# reordering DF steps:
tableFreqO=tableFreq[order(tableFreq$pct),]


base = ggplot(tableFreqO, aes(locale,pct)) 
lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = locale, 
                                   yend = pct, 
                                   xend = locale), color = "gray") 
lolliplot2 = lolliplot1 + geom_point()
lolliplot2 + scale_x_discrete(limits=tableFreqO$locale) # key element
```

And, what about changing the axis values so that we can identify the gaps:

```{r lolli_axis_gg, eval=FALSE}
# new variable
tableFreqO$gap=tableFreqO$pct-25

# plot the new variable
base = ggplot(tableFreqO, aes(locale,gap)) 

lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = locale, 
                                   yend = gap, 
                                   xend = locale), color = "gray") 
lolliplot2 = lolliplot1 + geom_point()
lolliplot2 + scale_x_discrete(limits=tableFreqO$locale) # key element

```

Maybe add some color:
```{r lolli_gg_color, eval=FALSE}
# a new column for color
tableFreqO$PositiveGap=ifelse(tableFreqO$gap>0,T,F)

# add new aesthetics 'color'
base = ggplot(tableFreqO, aes(locale,gap,
                              color=PositiveGap)) #change
lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = locale, 
                                   yend = gap, 
                                   xend = locale), color = "gray") 
lolliplot2 = lolliplot1 + geom_point()
lolliplot2 + scale_x_discrete(limits=tableFreqO$locale) # key element
```

Maybe add some extra info:
```{r lolli_annot_gg, eval=FALSE}
# a new column for color
tableFreqO$PositiveGap=ifelse(tableFreqO$gap>0,T,F)

base = ggplot(tableFreqO, aes(locale,gap,color=PositiveGap,
                              label = round(gap,3))) #  change
lolliplot1=base + geom_segment(aes(y = 0, 
                                   x = locale, 
                                   yend = gap, 
                                   xend = locale), color = "gray") 
lolliplot2=lolliplot1 + geom_point() 
lolliplot3= lolliplot2 + scale_x_discrete(limits=tableFreqO$locale) 
# annotating and moving the text on the horizontal
lolliplot3 + geom_text(nudge_x=0.3) 

```

You can avoid the overlaping symbols in the legend by using:
```{r nudge_gg, eval=FALSE}
lolliplot3 + geom_text(nudge_x=0.3,show.legend = FALSE) 
```

<span style="color:red"> Exercise:<br>Complete adding the elements missing in the last plot. </span>